Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-04 study_workload_1.tex
change some words in Section 4

The study of parallel workload~\Rmnum{1} consists of two parts. Firstly, we analyze the over-all network performance when Workload~\Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each application from the workload, analyze its performance by scrutinizing the traffic going through the routers that serve each job. The analysis allows us to identify the ``bully" in the workload.

@@

We study the network performance under different placement and routing configurations by analyzing the traffic going through the routers and their saturated time. 
The distribution of traffic over the network can reveal the loading situation, 
indicating whether the network is load-balanced or not. 
The saturated time can indicate the congestion situation on certain part of the network.

@@

Figure~\ref{fig:wkld-network-traffic-stime} shows the aggregate traffic for terminal links, local and global channels 
as well as the corresponding saturated time when 
Workload~\Rmnum{1} is running on the dragonfly network with six different placement and routing configurations. 
When contiguous placement coupled with minimal routing(CM) is in use, 
all the traffic are confined within the contiguously allocated groups, 
causing congestion on some routers along certain shortest paths from the source to the destination. 
The congested routers are over-loaded with high volume of traffic going through their local and global channels, 
shown as the blue dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. 
In the meanwhile, the corresponding saturated time of their local and global channels are also the highest compared with other configurations, 
shown as the blue dashed line in Figure~\ref{fig:global-channel-stime} and Figure~\ref{fig:local-channel-stime} .

When contiguous placement coupled with progressive adaptive routing (CPA) is in use, 
the traffic would take non-shortest paths via intermediate routers to void congestion. 
The congested routers along the shortest paths will not suffer the high volume of traffic as they used to. 
The traffic through local and global channels are greatly reduced, 
shown as the green dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. 
The corresponding saturated time on local and global channels also drop a lot, 
indicating the congestion are alleviated by using progressive adaptive routing. 
The adaptive routing performs comparably compared with progressive adaptive routing, 
thus CA is overlapped with CPA in the all the figures above.

Random placement policy performs comparably with three different routing policies.
Random placement can uniformly distribute MPI ranks over the network, 
balance the traffic load over the network. 
Shown as the solid lines in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}, 
no router suffers exceptional high volume of traffic on its local and global channels. 
When the random placements coupled with minimal routing (RM), 
the packets can avoid unnecessary intermediate forwarding as introduced by (progressive) adaptive routing, thus generating less traffic. 
The blue solid line(RM) is lower than the red and green solid line (RA, RPA) in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}.

Due to the same reason, 
no router suffers exceptionally long saturated time on their global channels when random placement coupled (progressive) adaptive are in use, 
shown as red and green solid lines in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. 
However, when random placement is couple with minimal routing, 
the saturated time on certain routers along the shortest paths would increase, 
shown as the prolonged blue solid line in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. 
Since the minimal routing always chooses the shortest paths, 
increasing the possibility of congestion along those paths.

%terminal link traffic and busy time explanation
Figure~\ref{fig:terminal-link-traffic} is for the purpose of symmetry, 
it shows the traffic per rank (aka traffic through terminal links), 
which stay the same when different placement and routing configurations are in use. 
All six lines are overlapped in Figure~\ref{fig:terminal-link-traffic}. 
However, due to the traffic changes on local and channels, 
the saturated time of terminal links will experience variability when different routing policies are in use, as shown in Figure~\ref{fig:terminal-link-stime}. 
When contiguous placement coupled with minimal routing(CM), 
the accumulated congestion on local and global channels cause extremely long saturated time on the terminal links, 
shown as the blue dashed line in Figure~\ref{fig:terminal-link-stime}. 
This again manifests the severe congestion caused by using contiguous placement and minimal routing.

@@

We study the performance of each application individually by isolating it from Workload~\Rmnum{1}.   
Figure~\ref{fig:apps-commtime} shows the distribution of communication time of three applications in per rank manner, 
when they running concurrently under six different placement and routing configurations listed in Table\ref{tab: placement routing configs}.

As shown in Figure~\ref{fig:mg-commtime}, 
MultiGrid takes the most communication time when it is running with contiguous placement and minimal routing (CM). 
When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA) are in use, 
the communication time is significantly reduced. 
Random placement coupled with minimal routing (RM) can make further improvement, 
and reaches the best performance(lowest communication time) when coupled with adaptive and progressive adaptive routing(RA, RPA). 
Due to the variance of data transfer amount among MPI ranks, 
MultiGrid communication time also presents great variation, 
indicating by the long boxes in Figure~\ref{fig:mg-commtime}.

>>>>

The study of Workload~\Rmnum{1} consists of two parts. Firstly, we analyze the overall network performance when Workload~\Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each application from the workload, analyze its performance by scrutinizing the traffic going through its serving routers. The analysis allows us to identify the ``bully" in Workload~\Rmnum{1}.

@@

We study the network performance by analyzing the traffic going through all the routers in the network and their saturated time. 
The traffic distribution can reveal the loading situation in the network, 
indicating whether the network is load-balanced or not. 
The saturated time can indicate the congestion situation on certain part of the network.

@@

Figure~\ref{fig:wkld-network-traffic-stime} shows the aggregate traffic for terminal links, local and global channels 
as well as the corresponding saturated time when 
Workload~\Rmnum{1} is running on the dragonfly network with six different placement and routing configurations. 
%explain CM
When contiguous placement coupled with minimal routing(CM) is in use, 
all the traffic are confined within the consecutively allocated groups, 
causing congestion on some routers along certain shortest paths from the source to the destination. 
The congested routers are over-loaded with high volume of traffic going through their local and global channels, 
shown as the blue dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. 
In the meanwhile, the corresponding saturated time of their local and global channels are also the highest compared with other configurations, 
as shown in Figure~\ref{fig:global-channel-stime} and Figure~\ref{fig:local-channel-stime} .
%explain CA and CPA
When contiguous placement coupled with progressive adaptive routing (CPA) is in use, 
the traffic would take non-shortest paths via intermediate routers to void congestion. 
The congested routers along the shortest paths will not suffer the high volume of traffic as they used to be. 
The traffic through local and global channels are greatly reduced, 
shown as the green dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. 
The corresponding saturated time on local and global channels also drop a lot, 
indicating the congestion are alleviated by using progressive adaptive routing. 
The adaptive routing performs comparably compared with progressive adaptive routing, 
thus CA is overlapped with CPA in the all the figures above.

%Explain RM
Random placement policy performs comparably with three different routing policies.
Random placement can uniformly distribute MPI ranks over the network, 
balance the traffic load over the network. 
Shown as the solid lines in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}, 
no router suffers exceptional high volume of traffic on its local and global channels. 
When the random placements coupled with minimal routing (RM), 
the packets can avoid unnecessary intermediate forwarding as introduced by (progressive) adaptive routing, thus generating less traffic. 
That's why blue solid line(RM) is lower than the red and green solid line (RA, RPA) in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. 
%Explain RA, RPA
For the same reason, 
no router suffers exceptionally long saturated time on their global channels when random placement coupled (progressive) adaptive are in use, 
shown as red and green solid lines in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. 
However, when random placement is couple with minimal routing, 
the saturated time on certain routers along the shortest paths would increase, 
shown as the prolonged blue solid line in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. 
Since the minimal routing always chooses the shortest paths, 
increasing the possibility of congestion along those paths.

% Explain terminal link traffic and busy time
Figure~\ref{fig:terminal-link-traffic} is for the purpose of symmetry, 
it shows the traffic per rank (aka traffic through terminal links), 
which stay the same when different placement and routing configurations are in use. 
All six lines are overlapped in Figure~\ref{fig:terminal-link-traffic}. 
However, due to the traffic changes on local and channels, 
the saturated time of terminal links will experience variability when different routing policies are in use, as shown in Figure~\ref{fig:terminal-link-stime}. 
When contiguous placement coupled with minimal routing(CM), 
the accumulated congestion on local and global channels cause extremely long saturated time on the terminal links, 
shown as the blue dashed line in Figure~\ref{fig:terminal-link-stime}. 
This again manifests the severe congestion caused by using contiguous placement and minimal routing.

@@

We study the performance of each application individually by isolating it from Workload~\Rmnum{1}.   
Figure~\ref{fig:apps-commtime} shows the distribution of their communication time when running concurrently under different placement and routing configurations.
As shown in Figure~\ref{fig:mg-commtime}, 
MultiGrid takes the most communication time when it is running with contiguous placement and minimal routing (CM). 
When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA) are in use, 
the communication time is significantly reduced. 
Random placement coupled with minimal routing (RM) can make further improvement, 
and reaches the best performance(lowest communication time) when coupled with adaptive and progressive adaptive routing(RA, RPA). 
Due to the variance of data transfer amount among MPI ranks, 
MultiGrid communication time also presents great variation, 
indicating by the long boxes in Figure~\ref{fig:mg-commtime}.
